\sout{Machine Learning for Global Conquest} \\ Multimatrices and Their Derivatives

### 2020-08-06 Math/book.tex
About to update computer, saving work.

Now, not knowing what else to do with the results, I've written this book
about my amature exploration into multidimensional matricies -
multimatricies - and how they can simplify derivatives of matrix equations.

@@

Let one matrix $X$ represent the level of scientific education level of a set of
undergraduates who work in a densely packed 3 by 5 office.
Then let $B$ represent the belief each of those undergradutes has in a higher power.

@@

Suppose the shape of these two matricies are defined,
$\shape{B} = \shape{X} = \left< 3,5 \right>$.
And each element of $B$ will be the sine of the corresponding element of
$X$. That is, \[B_{r,c} = \sin(X_{r,c})\]
Which I'll write more simply as, \[B = \sin(X)\]

Now, I would call the derivative of any element of $B$ with respect to any element
of $X$ a partial derivative of that function. Since
$\partial B_{r_1, c_1}/\partial X_{r_2, c_2}$ is defined as the rate of change of
\textit{one} undergraduate's belief in a higher power with respect to \textit{one}
undergraduate's education level - it only represents a change to part of the whole
system. 
We can calculate this partial derivative easily enough,
\[
\frac{\partial B_{r_1,c_1}}{\partial X_{r_2,c_2}} = 
\left\{
  \begin{array}{ll}
    \cos(X_{r_2,c_2})  & \mbox{if } r_1 = r_2 \mbox{ and } c_1 = c_2 \\
    0 & \mbox{otherwise}
  \end{array}
\right.
\]
Every student's belief in a higher power changes with respect to their own
scientific education level but does not change with respect to other students'
scientific education. No mater how much I teach Jane about photosynthesis, I'm
not going to change Joe's mind about divine mysteries.

@@

So what is this object we've defined? In all ways that matter, it seems to be
a multidimensional matrix. It has four dimensions of indices, rather than just
a row and column but each uniquely indexed cell in this object may have a
unique value, just like a standard two dimensional matrix. Let's define this
bit of notation formaly.

@@

Now that you seem to understand.. Don't give me that look... well if you're confused
you should have been asking questions, moving on. Now that \textit{most of you}
seem to have understood, I'll give you a theorem to prove.

@@

Going back to our original motivating problem, we (that is to say, myself and my
split personalities whom I reference as to give my writing the percieved authority
of more than one author) can express our derivative notation thusly,

@@

The multiplication of traditional matricies is a calculation that, when first
encoutered, inspires thoughts of omnicide.

@@

But even when the dimensional analysis makes sense, multimatrix multiplication
may not be associative (Honestly, how could you be so naive?). 
Take, for example, multimatricies $A$, $B$, and $C$ such
that $\shape{A} = \left<2,2,2,2\right>$, $\shape{B} = \left<2,2,2\right>$, and
$\shape{C} = \left<2,2,2,2\right>$. Let,
\[
 A[\bar{a}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{a} = \left<1,1,1,2\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
\[
 B[\bar{b}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{b} = \left<1,2,1\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
And
\[
 C[\bar{c}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{c} = \left<1,1,1,1\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
Then,
\[
 (A \mmult{2} B)[\bar{x}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{x} = \left<1,1,1\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
So,
\[
 ((A \mmult{2} B) \mmult{2} C)[\bar{x}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{x} = \left<1,1\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
But, $(B \mmult{2} C)[\bar{x}] = 0$ so $(A \mmult{2} (B \mmult{2} C))[\bar{x}] = 0$
Which goes to show that even though both $(A \mmult{2} B) \mmult{2} C$ and
$A \mmult{2} (B \mmult{2} C)$ are valid calculations which result in multimatrixes
of the same dimensionality, they are not equal so multiplicative associativity
doesn't hold. It also goes to show that your intuition is bad and you should feel bad.

@@

\begin{theorem}[Communicativity of Multiplication]
Let $\shape{A} = \vec{x}$ and $\shape{B} = \vec{x}$. Then,
\[ A \mmult{\shape{\vec{x}}} B = B \mmult{\shape{\vec{x}}} A \]
\end{theorem}
\begin{proof}
Let $A$, $B$, and $\vec{x}$ be as described above. Notice that then $\shape{A \mmult{\shape{\vec{x}}} B}$
results in a zero dimensional multimatrix - a multimatrix with one element which is effectively
just a scalar.
\begin{align*}
(A \mmult{\shape{\vec{x}}} B)[\{\}]
  &= \sum_{\forall \bar{x} \in \Dim(\vec{x})} A[\bar{x}]B[\bar{x}] \\
  &= \sum_{\forall \bar{x} \in \Dim(\vec{x})} B[\bar{x}]A[\bar{x}] \\
  &= (B \mmult{\shape{\vec{x}}} A)[\{\}]
\end{align*}
\end{proof}

@@

\begin{definition}[Multimatrix Identities]
\label{mm_mult_ident}
Let us define $\Ident^n(\vec{v})$ as the $n^{\text{th}}$ order identity with base shape
$\vec{v}$. This $\Ident^n(\vec{v})$ is a multimatrix with shape,
\[ \shape{\Ident^n(\vec{v})} = \bigoplus_{i \in [1, n]} \vec{v} \]
That is, it's shape is $n$ successive concatinations of $\vec{v}$.
Let each element of $\Ident^n(\vec{v})$ is defined as,
\[ \forall \bar{v_1}, \bar{v_2}, \ldots, \bar{v_n} \text{ each in } \Dim(\vec{v}) : \]
\[
 \Ident^n(\vec{v})[\bar{v_1}, \bar{v_2}, \ldots, \bar{v_n}]
 = \left\{
  \begin{array}{ll}
    1 & \mbox{if } \bar{v_1} = \bar{v_2} = \ldots = \bar{v_n} \\
    0 & \mbox{otherwise}
  \end{array}
 \right.
\]

@@

That definition seems rather arbitrary but stick with me, I know what I'm
talking about. I've got a \sout{doctorate}
masters degree in \sout{mathematics} I mean \sout{a real engineering field}, fine, software
engineering. Whatever. Just look at how the
square identity it relates to a multiplication operation.

@@

@@

The cubic identity, strangely enough, appear in the derivative of elementwise
functions. Which is one of the reaspons I needed to expand my definition of
identities beyond $\Ident^2(\vec{v})$.

\begin{theorem}[Elementwise Derivative]
\label{elementwise_derivative}
Let $M$ be an element wise function on $X$. That is, Let $M(X)$ be a function
such that,
\[ M(X) : \mathbb{R}^{\vec{x}} \to \mathbb{R}^{\vec{x}} \]
And which applies the same scalar function to each element of $X$, 
\[ M(X)[\bar{x}] = m(X[\bar{x}]) \]
Also, let $M'$ be the elementwise function on $X$ which applies the derivative
of $M$'s associated scalar function to each element of $X$. That is,
\[ M'(X)[\bar{x}] = m'(X[\bar{x}]) \]
Then,
\[ \frac{dM}{dX} = \Ident^3(\shape{X}) \mmult{\shape{\shape{X}}} M'(X) \]
\end{theorem}
\begin{proof}
By the definition of a multimatrix derivative,
\[ \forall \bar{m} \in \Dim(\shape{M}), \bar{x} \in \Dim(\shape{X}) : \]
\begin{align*}
 \frac{dM}{dX}[\bar{m} \oplus \bar{x}]
 &= \frac{\partial M[\bar{m}]}{\partial X[\bar{x}]} \\
 &= \frac{\partial m(X[\bar{m}])}{\partial X[\bar{x}]} \\
 &= \left\{
  \begin{array}{ll}
    m'(X[\bar{m}]) & \mbox{if } \bar{m} = \bar{x} \\
    0 & \mbox{otherwise}
  \end{array}
 \right. \\
 &= \delta_{\bar{m} \bar{x}} m'(X[\bar{m}])
\end{align*}
Since, by our definition of elementwise function, $\shape{M} = \shape{X}$, we can
say that, 
\begin{align*}
 (\Ident^3(\shape{X}) \mmult{\shape{\shape{X}}} M'(X))[\bar{m} \oplus \bar{x}]
 &= \sum_{\forall \bar{y} \in \Dim(\shape{X})}
 \Ident^3(\shape{X})[\bar{m} \oplus \bar{x} \oplus \bar{y}] M'(X)[\bar{y}] \\
 &= \sum_{\forall \bar{y} \in \Dim(\shape{X})}
 \Ident^3(\shape{X})[\bar{m} \oplus \bar{x} \oplus \bar{y}] m'(X[\bar{y}]) \\
 &= \sum_{\forall \bar{y} = \bar{m} = \bar{x}}
 \Ident^3(\shape{X})[\bar{m} \oplus \bar{x} \oplus \bar{y}] m'(X[\bar{y}]) \\
 &= \left\{
  \begin{array}{ll}
    m'(X[\bar{m}]) & \mbox{if } \bar{m} = \bar{x} \\
    0 & \mbox{otherwise}
  \end{array}
 \right. \\
 &= \delta_{\bar{m} \bar{x}} m'(X[\bar{m}])
\end{align*}
Therefore $dM/dX = \Ident^3(\shape{X}) \mmult{\shape{\shape{X}}} M'(X)$
\end{proof}

@@

\begin{landscape}
\begin{theorem}[Identity Contraction]
For all $m, n, k \in \mathbb{N}$ (note $m$, $n$, and $k$ are non-zero) and $\vec{x}$ as a valid multimatrix shape,
where $m > k$ and $n > k$,
\[ \Ident^m(\vec{x}) \mmult{k\shape{\vec{x}}} \Ident^n(\vec{x}) = \Ident^{m+n-2k}(\vec{x}) \]
\end{theorem}
\begin{proof}
Consider that for some indicies
$\{\bar{a_1}, \bar{a_2}, \ldots \bar{a}_{m-k}\}$
and $\{\bar{b_1}, \bar{b_2}, \ldots \bar{b}_{n-k}\}$ where each
$\bar{a_i} \in \Dim(\vec{x})$ and each $\bar{b_i} \in \Dim(\vec{x})$,

\begin{align*}
&\left( \Ident^m(\vec{x}) \mmult{k\shape{\vec{x}}} \Ident^n(\vec{x}) \right)
[\bar{a_1} \oplus \bar{a_2} \oplus \ldots \bar{b_1} \oplus \bar{b_2} \oplus \ldots] = \\
&\sum_{\forall \bar{c_1}, \bar{c_2}, \ldots \bar{c_k} \in \Dim(\vec{x})}
\Ident^m(\vec{x})
[\bar{a_1} \oplus \bar{a_2} \oplus \ldots \bar{c_1} \oplus \bar{c_2} \oplus \ldots]
\Ident^n(\vec{x})
[\bar{c_1} \oplus \bar{c_2} \oplus \ldots \bar{b_1} \oplus \bar{b_2} \oplus \ldots]
\end{align*}

Consider that $\Ident^m(\vec{x})
[\bar{a_1} \oplus \bar{a_2} \oplus \ldots \bar{c_1} \oplus \bar{c_2} \oplus \ldots]$
will be 0 whenever the any $\bar{c_i} \ne \bar{c_j}$ - we can simplify that to all
the same $c$.

\begin{align*}
&\left( \Ident^m(\vec{x}) \mmult{k\shape{\vec{x}}} \Ident^n(\vec{x}) \right)
[\bar{a_1} \oplus \bar{a_2} \oplus \ldots \bar{b_1} \oplus \bar{b_2} \oplus \ldots] = \\
&\sum_{\forall \bar{c} \in \Dim(\vec{x})}
\Ident^m(\vec{x})
[\bar{a_1} \oplus \bar{a_2} \oplus \ldots \bar{c} \oplus \bar{c} \oplus \ldots]
\Ident^n(\vec{x})
[\bar{c} \oplus \bar{c} \oplus \ldots \bar{b_1} \oplus \bar{b_2} \oplus \ldots]
\end{align*}

Consider also that there can be at most one $\bar{c}$ for which
$\bar{c} = \bar{a_1} = \bar{a_2} \ldots$. And since our right hand side is
the sum of products of ones and zeros, we can therefore conclude it is either 1 or 0.

We can futher reason that the right hand side can only be one when all of the
$\bar{a_i}$ and all of the $\bar{b_i}$ equal $\bar{c}$. Furthermore, if all
$\bar{a_i}$ and all $\bar{b_i}$ are equal then that $\bar{c}$ must exist since it can
be any value in the range of all $\bar{a_i}$ and $\bar{b_i}$ and this is not a
zero width multiplication ($k >= 1$). Therefore,

\begin{align*}
\left( \Ident^m(\vec{x}) \mmult{k\shape{\vec{x}}} \Ident^n(\vec{x}) \right)
[\bar{a_1} \oplus \bar{a_2} \oplus \ldots \bar{b_1} \oplus \bar{b_2} \oplus \ldots] =
\left\{
  \begin{array}{ll}
    1 & \mbox{if } \bar{a_1} = \bar{a_2} = \ldots = \bar{b_1} = \bar{b_2} = \ldots \\
    0 & \mbox{otherwise}
  \end{array}
\right.
\end{align*}

Which is an identity matrix! And since the dimensional rules of multimatrix 
multiplication show that,

\[
 \shape{\shape{ \Ident^m(\vec{x}) \mmult{k\shape{\vec{x}}} \Ident^n(\vec{x}) }}
 =
 \shape{\shape{\Ident^m(\vec{x})}} + \shape{\shape{\Ident^n(\vec{x})}} - 2k\shape{\vec{x}}
\]

We've shown that,

>>>>

Now, not knowing what else to do with the results, I've written this book
about my exploration into multidimensional matricies,
multimatricies, and how they can elucidate derivatives of matrix equations.

@@

Let's say we have a small class of undergraduates packed, conveniently in a 3 by 5
grid. We'll let $X$ be the matrix representing number of semesters of science
classes each of those student has taken.
Then let $B$ represent the belief each of those undergradutes has in a higher power,
thier BHP if you will.

@@

Let's formally state the shape of these two matricies,
$\shape{B} = \shape{X} = \left< 3,5 \right>$. (Throughout this work, I'll be using
$\shape{M}$ to denote the shape of $M$ rather than the determinate because, frankly, it
is just comes up more frequently.)
Each element of $B$ will be the sine of the corresponding element of
$X$. That is, \[B_{r,c} = \sin(X_{r,c})\]
Which I'll write more simply as, \[B = \sin(X)\]

Now, I would call the derivative of any element of $B$ with respect to any element
of $X$ a partial derivative of that function. Since
$\partial B_{r_1, c_1}/\partial X_{r_2, c_2}$ is defined as the rate of change of
\textit{one} undergraduate's BPH with respect to \textit{one}
undergraduate's education level - it only represents a change to part of the whole
system. 
We can calculate this partial derivative easily enough,
\[
\frac{\partial B_{r_1,c_1}}{\partial X_{r_2,c_2}} = 
\left\{
  \begin{array}{ll}
    \cos(X_{r_2,c_2})  & \mbox{if } r_1 = r_2 \mbox{ and } c_1 = c_2 \\
    0 & \mbox{otherwise}
  \end{array}
\right.
\]
Every student's BHP changes with respect to their own
scientific education level but does not change with respect to other students'
scientific education. No mater how much I teach Jane about photosynthesis, I'm
not going to change Joe's mind about divine mysteries.

@@

So what is this object we've defined? In all ways that matter, it seems to be
a multidimensional matrix. It has four dimensions of indices, rather than just
a row and column but each uniquely indexed cell in this object may have a
unique value, just like a standard two dimensional matrix.

Programmers might call this object a `tensor' but mathematitians might
scream at those same programmers, ``WE ALREADY HAVE A DEFINITION OF A TENSOR! DIE!''
And the ensuing war might wipe out most of the eastern seaboard and would
certainly not be conductive to a reasonable discussion. To avoid such a
confrontation I'll be using the incredibly imaginative term `multimatrix'.

Let's define this bit of notation formaly.

@@

\begin{definition}[Scalar Filled Multimatrices]
For convenience, I'll often refer to multimatricies filled with a single value,
$s \in \mathbb{R}$,
using the notation $s^{\vec{x}}$ where $\vec{x}$ is some multimatrix shape. So that
$\shape{s^{\vec{x}}} = \vec{x}$ and for all $\bar{x} \in \Dim(\vec{x})$,
\[ s^{\vec{x}}[\bar{x}] = s \]

For example, $0^{\left<2,3,2\right>}$ is a multimatrix with shape $\left<2,3,2\right>$
that is filled with zeros.
\end{definition}

Now that you seem to understand.. Don't give me that look... well if you're confused
you should have been asking questions. I don't have time to backtrack now.
Well, seeing as \textit{most of you} seem to have understood, I'll give you a theorem to prove.

@@

Going back to our original motivating problem, we (that is to say, myself and the voices
in my head whom I reference as to give my writing the percieved authority
of more than one author) can express our derivative notation thusly,

@@

The multiplication of traditional matricies is often considered a rather confusing
proposition, especially when one first encounters it.

@@

But even when the dimensional analysis makes sense, multimatrix multiplication
may not be associative (Honestly, how could you be so naive?). 
Take, for example, multimatricies $A$, $B$, and $C$ such
that $\shape{A} = \left<2,2,2,2\right>$, $\shape{B} = \left<2,2,2\right>$, and
$\shape{C} = \left<2,2,2,2\right>$. Let,
\[
 A[\bar{a}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{a} = \left<1,1,1,2\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
\[
 B[\bar{b}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{b} = \left<1,2,1\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
And
\[
 C[\bar{c}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{c} = \left<1,1,1,1\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
Then,
\[
 (A \mmult{2} B)[\bar{x}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{x} = \left<1,1,1\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
So,
\[
 ((A \mmult{2} B) \mmult{2} C)[\bar{x}] = 
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \bar{x} = \left<1,1\right> \\
      0 & \mbox{otherwise}
    \end{array}
  \right.
\]
But, $B \mmult{2} C = 0^{\left<2,2,2\right>}$ so $A \mmult{2} (B \mmult{2} C) = 0^{\left<2,2,2\right>}$
Which goes to show that even though both $(A \mmult{2} B) \mmult{2} C$ and
$A \mmult{2} (B \mmult{2} C)$ are valid calculations which result in multimatrixes
of the same dimensionality, they are not equal so multiplicative associativity
doesn't hold. It also goes to show that your intuition is bad and you should feel bad.

@@

\begin{theorem}[Communicativity of Multiplication]
Let $\shape{A} = \vec{x}$ and $\shape{B} = \vec{x}$. Then,
\[ A \mmult{\shape{\vec{x}}} B = B \mmult{\shape{\vec{x}}} A \]
\end{theorem}
\begin{proof}
Let $A$, $B$, and $\vec{x}$ be as described above. Notice that then $\shape{A \mmult{\shape{\vec{x}}} B}$
results in a zero dimensional multimatrix - a multimatrix with one element which is just a scalar.
\begin{align*}
(A \mmult{\shape{\vec{x}}} B)[\left<\right>]
  &= \sum_{\forall \bar{x} \in \Dim(\vec{x})} A[\bar{x}]B[\bar{x}] \\
  &= \sum_{\forall \bar{x} \in \Dim(\vec{x})} B[\bar{x}]A[\bar{x}] \\
  &= (B \mmult{\shape{\vec{x}}} A)[\left<\right>]
\end{align*}
\end{proof}

@@

\begin{definition}[Multimatrix Identities]
\label{mm_mult_ident}
Let us define $\Ident^n(\vec{v})$ as the $n^{\text{th}}$ order identity with base shape
$\vec{v}$. This $\Ident^n(\vec{v})$ is a multimatrix with shape,
\[ \shape{\Ident^n(\vec{v})} = \bigoplus_{i = 1}^n \vec{v}  = n * \vec{v} \]
That is, it's shape is $n$ successive concatinations of $\vec{v}$.
Let each element of $\Ident^n(\vec{v})$ is defined as,
\[ \forall \bar{v_1}, \bar{v_2}, \ldots, \bar{v_n} \text{ each in } \Dim(\vec{v}) \]
\[ \bigoplus_{i = 1}^n \bar{v_i} = \bar{v_I} \]
\[
 \Ident^n(\vec{v})[\bar{v_I}]
 = \left\{
  \begin{array}{ll}
    1 & \mbox{if } \bar{v_1} = \bar{v_2} = \ldots = \bar{v_n} \\
    0 & \mbox{otherwise}
  \end{array}
 \right.
\]

@@

That definition seems rather arbitrary but stick with me, I know what I'm
talking about. What credentials do I have? Totally irrelevant. Stop asking questions.

Now, please direct your attention to this example of how the square identity relates
to a multiplication operation.

@@

Multimatrix identities also have the nice property of being communicative when multiplied
across multiples of their base shape and completely engulfing a smaller multimatrix.

\begin{theorem}[Communicativity of Multiplicative Identities]
Let $X$ be a multimatrix of shape $k * \vec{x}$ where $k$ is some natural number.
Remember (from the identity definition) that the $*$ operator denote successive concatenations of a vector so that,
\[ k * \vec{x} = \bigoplus_{i=1}^k \vec{x} \]
Then for any $n \ge k$,
\[ X \mmult{k \shape{\vec{x}}} \Ident^n(\vec{x}) = \Ident^n(\vec{x}) \mmult{k \shape{\vec{x}}} X \]
\end{theorem}
\begin{proof}
Given that $\shape{X} = k * \vec{x}$ we know that
$\shape{X \mmult{k \shape{\vec{x}}} \Ident^n(\vec{x})} = (n-k) * \vec{x}$.
Let $\bar{x}_1, \bar{x}_2 \ldots \bar{x}_{n-k}$ each be index vectors in $\Dim(\vec{x})$,
for simplicity denoted as,
\[ \bar{x}_I = \bigoplus_{i=1}^{n-k} \bar{x}_i \]
We'll also need a similarly denoted $\bar{y}_1 \ldots \bar{y}_k$ for the free indicies of
the sum below and that shall also be represented as,
\[ \bar{y}_J = \bigoplus_{j=1}^{k} \bar{y}_j \]
Then we know that,
\[
  \left(X \mmult{k \shape{\vec{x}}} \Ident^n(\vec{x}) \right)[\bar{x}_I]
  = \sum_{\forall \bar{y}_J} X[\bar{y}_J] \Ident^n(\vec{x})[\bar{y}_J \oplus \bar{x}_I]
\]
Notice that $\Ident^n(\vec{x})[\bar{y}_J \oplus \bar{x}_I] = \Ident^n(\vec{x})[\bar{x}_I \oplus \bar{y}_J]$
since the sub index vectors will always either be equal or not equal, $1$ or $0$, their order doesnt matter.
So,
\begin{align*}
  \left(X \mmult{k \shape{\vec{x}}} \Ident^n(\vec{x}) \right)[\bar{x}_I]
  &= \sum_{\forall \bar{y}_J} X[\bar{y}_J] \Ident^n(\vec{x})[\bar{x}_I \oplus \bar{y}_J] \\
  &= \sum_{\forall \bar{y}_J} \Ident^n(\vec{x})[\bar{x}_I \oplus \bar{y}_J] X[\bar{y}_J] \\
  &= \left(\Ident^n(\vec{x}) \mmult{k \shape{\vec{x}}} X \right)[\bar{x}_I]
\end{align*}
\end{proof}

@@

The cubic identity, strangely enough, appear in the derivative of elementwise
functions. Which is one of the reasons I needed to expand my definition of
identities beyond $\Ident^2(\vec{v})$.

\begin{theorem}[Elementwise Derivative]
\label{elementwise_derivative}
Let $M$ be an element wise function on $X$. That is, Let $M(X)$ be a function
such that,
\[ M(X) : \mathbb{R}^{\vec{x}} \to \mathbb{R}^{\vec{x}} \]
And which applies the same scalar function to each element of $X$, 
\[ M(X)[\bar{x}] = m(X[\bar{x}]) \]
Also, let $M'$ be the elementwise function on $X$ which applies the derivative
of $M$'s associated scalar function to each element of $X$. That is,
\[ M'(X)[\bar{x}] = m'(X[\bar{x}]) \]
Then,
\[ \frac{dM}{dX} = \Ident^3(\shape{X}) \mmult{\shape{\shape{X}}} M'(X) \]
\end{theorem}
\begin{proof}
By the definition of a multimatrix derivative,
\[ \forall \bar{m}, \bar{x} \in \Dim(\shape{X}) : \]
\begin{align*}
 \frac{dM}{dX}[\bar{m} \oplus \bar{x}]
 &= \frac{\partial M[\bar{m}]}{\partial X[\bar{x}]} \\
 &= \frac{\partial m(X[\bar{m}])}{\partial X[\bar{x}]} \\
 &= \left\{
  \begin{array}{ll}
    m'(X[\bar{m}]) & \mbox{if } \bar{m} = \bar{x} \\
    0 & \mbox{otherwise}
  \end{array}
 \right. \\
 &= \delta_{\bar{m} \bar{x}} m'(X[\bar{m}])
\end{align*}
Since, by our definition of elementwise function, $\shape{M} = \shape{X}$, we can
say that, 
\begin{align*}
 (\Ident^3(\shape{X}) \mmult{\shape{\shape{X}}} M'(X))[\bar{m} \oplus \bar{x}]
 &= \sum_{\forall \bar{y} \in \Dim(\shape{X})}
 \Ident^3(\shape{X})[\bar{m} \oplus \bar{x} \oplus \bar{y}] M'(X)[\bar{y}] \\
 &= \sum_{\forall \bar{y} \in \Dim(\shape{X})}
 \Ident^3(\shape{X})[\bar{m} \oplus \bar{x} \oplus \bar{y}] m'(X[\bar{y}]) \\
 &= \sum_{\forall \bar{y} = \bar{m} = \bar{x}}
 \Ident^3(\shape{X})[\bar{m} \oplus \bar{x} \oplus \bar{y}] m'(X[\bar{y}]) \\
 &= \left\{
  \begin{array}{ll}
    m'(X[\bar{m}]) & \mbox{if } \bar{m} = \bar{x} \\
    0 & \mbox{otherwise}
  \end{array}
 \right. \\
 &= \delta_{\bar{m} \bar{x}} m'(X[\bar{m}])
\end{align*}
Therefore $dM/dX = \Ident^3(\shape{X}) \mmult{\shape{\shape{X}}} M'(X)$
\end{proof}

Another important fact about the identity multimatrix is that it can be
reduced by multiplication with a $1^{\vec{s}}$ matrix.

\begin{theorem}[Identity Reduction]
Let $\vec{x}$ be a valid multimatrix shape and $m$ and $n$ be natural numbers
such that $m < n$.
Remember (from the identity definition) that the $*$ operator denote successive
concatenations of a vector so that,
\[ m*\vec{x} = \bigoplus_{i=1}^{m} \vec{x} \]
Then,
\[1^{m*\vec{x}} \mmult{m \shape{\vec{x}}} \Ident^n(\vec{x}) = \Ident^{n-m}(\vec{x})\]
and,
\[\Ident^n(\vec{x}) \mmult{m \shape{\vec{x}}} 1^{m*\vec{x}} = \Ident^{n-m}(\vec{x})\]
\end{theorem}
\begin{proof}
\begin{ppart}
Let $\bar{x}_1, \bar{x}_2 \ldots  \bar{x}_{n-m}$ each in $\Dim(\vec{x})$ and let
\[\bigoplus_{i = 1}^{n-m} \bar{x}_i = \bar{x}_I \]
In the sum below, let $\bar{y}_1, \bar{y}_2 \ldots \bar{y}_m$ each in $\Dim(\vec{x})$ be the
free indicies denoted,
\[\bigoplus_{j = 1}^m \bar{y}_j = \bar{y}_J \]
Then,
\begin{align*}
  1^{m*\vec{x}} \mmult{m \shape{\vec{x}}} \Ident^n(\vec{x}) [\bar{x}_I] &=
  \sum_{\forall \bar{y}_J} 1^{m*\vec{x}}[\bar{y}_J] \Ident^n(\vec{x})[\bar{y}_J \oplus \bar{x}_I] \\
  &=
  \sum_{\forall \bar{y}_J} (1) \Ident^n(\vec{x})[\bar{y}_J \oplus \bar{x}_I] \\
  &=
  \sum_{\forall \bar{y}_J} \Ident^n(\vec{x})[\bar{y}_J \oplus \bar{x}_I] \\
  &=
  \left(
    \sum_{\forall \bar{y}_J, \text{ each } \bar{y}_i \ne \bar{x}_1} \Ident^n(\vec{x})[\bar{y}_J \oplus \bar{x}_I]
  \right)
  + \Ident^n(\vec{x})\left[\left( \bigoplus_{j=1}^m \bar{x}_1 \right) \oplus \bar{x}_I\right] \\
  &= 0 + \Ident^n(\vec{x})\left[\left( \bigoplus_{j=1}^m \bar{x}_1 \right) \oplus \bar{x}_I\right] \\
  &= \Ident^{n-m}(\vec{x})[\bar{x}_I]
\end{align*}
\end{ppart}
\begin{ppart}
Let $\bar{x}_1, \bar{x}_2 \ldots \bar{x}_{n-m}$ each in $\Dim(\vec{x})$ and let
\[\bigoplus_{i = 1}^{n-m} \bar{x}_i = \bar{x}_I \]
In the sum below, let $\bar{y}_1, \bar{y}_2 \ldots \bar{y}_m$ each in $\Dim(\vec{x})$ be the
free indicies denoted,
\[\bigoplus_{j = 1}^m \bar{y}_j = \bar{y}_J \]
Then,
\begin{align*}
  \Ident^n(\vec{x}) \mmult{m \shape{\vec{x}}} 1^{m*\vec{x}} [\bar{x}_I] &=
  \sum_{\forall \bar{y}_J} \Ident^n(\vec{x})[\bar{x}_I \oplus \bar{y}_J] 1^{m*\vec{x}}[\bar{y}_J] \\
  &=
  \sum_{\forall \bar{y}_J} \Ident^n(\vec{x})[\bar{x}_I \oplus \bar{y}_J] (1) \\
  &=
  \sum_{\forall \bar{y}_J} \Ident^n(\vec{x})[\bar{x}_I \oplus \bar{y}_J] \\
  &=
  \left(
    \sum_{\forall \bar{y}_J, \text{ each } \bar{y}_i \ne \bar{x}_1} \Ident^n(\vec{x})[\bar{x}_I \oplus \bar{y}_J]
  \right)
  + \Ident^n(\vec{x})\left[\bar{x}_I \oplus \bigoplus_{j=1}^m \bar{x}_1 \right] \\
  &= 0 + \Ident^n(\vec{x})\left[\bar{x}_I \oplus \bigoplus_{j=1}^m \bar{x}_1 \right] \\
  &= \Ident^{n-m}(\vec{x})[\bar{x}_I]
\end{align*}
\end{ppart}
\end{proof}

@@

\begin{landscape}
\begin{theorem}[Identity Contraction]
For all $m, n, k \in \mathbb{N}$ (note $m$, $n$, and $k$ are non-zero) and $\vec{x}$ as a valid multimatrix shape,
where $m > k$ and $n > k$,
\[ \Ident^m(\vec{x}) \mmult{k\shape{\vec{x}}} \Ident^n(\vec{x}) = \Ident^{m+n-2k}(\vec{x}) \]
\end{theorem}
\begin{proof}
Consider that for some indicies vectors
$\bar{a}_1, \bar{a}_2 \ldots \bar{a}_{m-k}$
and $\bar{b}_1, \bar{b}_2 \ldots \bar{b}_{n-k}$ where each
$\bar{a_i} \in \Dim(\vec{x})$ and each $\bar{b_i} \in \Dim(\vec{x})$.
Now let,
\[ \bigoplus_{i=1}^{m-k} \bar{a}_i = \bar{a}_I \]
and,
\[ \bigoplus_{j=1}^{n-k} \bar{b}_j = \bar{b}_J \]
In the sums below we'll let $\bar{c}_1, \bar{c}_2 \ldots \bar{c}_k$ be free indicies
each in $\Dim(\vec{x})$ and denote,
\[ \bigoplus_{l = 1}^k \bar{c}_l = \bar{c}_K \]

Then,
\[
\left( \Ident^m(\vec{x}) \mmult{k\shape{\vec{x}}} \Ident^n(\vec{x}) \right)
[\bar{a}_I \oplus \bar{b}_J] = 
\sum_{\forall \bar{c}_K} \Ident^m(\vec{x}) [\bar{a}_I \oplus \bar{c}_K] \Ident^n(\vec{x}) [\bar{c}_K \oplus \bar{b}_J]
\]

Consider, for any $i$ if $\bar{a}_i \ne \bar{a}_1$ then
$\Ident^m(\vec{x}) [\bar{a}_I \oplus \bar{c}_K]$ will always be zero.
Similarly, if for any $j$, $\bar{b}_j \ne \bar{b}_1$ 
$\Ident^n(\vec{x}) [\bar{c}_K \oplus \bar{b}_J]$ will always be zero.

So, if we let, $\hat{a}_1 = (m-k) * \bar{a}_1$ and $\hat{b}_1 = (n-k) * \bar{b}_1$ then,
\begin{align*}
\left( \Ident^m(\vec{x}) \mmult{k\shape{\vec{x}}} \Ident^n(\vec{x}) \right) &
[\bar{a}_I \oplus \bar{b}_J] = \\
&\left\{
  \begin{array}{ll}
    \sum_{\forall \bar{c}_K}
      \Ident^m(\vec{x}) [\hat{a}_1 \oplus \bar{c}_K] 
      \Ident^n(\vec{x}) [\bar{c}_K \oplus \hat{b}_1]
    & \mbox{if } \forall i, j : \bar{a}_i = \bar{a}_1 \mbox{ and } \bar{b}_j = \bar{b}_1 \\
    0 & \mbox{otherwise}
  \end{array}
\right.
\end{align*}

Now, for any $\bar{c}_i$ that is not equal to $\bar{a}_1$, 
$\Ident^m(\vec{x}) [\hat{a}_1 \oplus \bar{c}_K]$ will be zero.
So, we only need to consider the part of the sum where $\bar{c}_K = k * \bar{a}_1$
\begin{align*}
\left( \Ident^m(\vec{x}) \mmult{k\shape{\vec{x}}} \Ident^n(\vec{x}) \right) &
[\bar{a}_I \oplus \bar{b}_J] = \\
&\left\{
  \begin{array}{ll}
    \Ident^m(\vec{x}) [\hat{a}_1 \oplus (k * \bar{a}_1)] 
    \Ident^n(\vec{x}) [(k * \bar{a}_1) \oplus \hat{b}_1]
    & \mbox{if } \forall i, j : \bar{a}_i = \bar{a}_1 \mbox{ and } \bar{b}_j = \bar{b}_1 \\
    0 & \mbox{otherwise}
  \end{array}
\right.
\end{align*}
